---
title: "EDS 222 Week 1 Wednesday"
subtitle: "Sampling"
format: html
---

We're going to see how to sample from both *empirical* and *parametric* distributions in R. This is the foundation of *simulation*, which is the most important tool in this course.

## Empirical Distributions

```{r}
library(tidyverse)
set.seed(222)
# Set all ggplots to theme_bw()
theme_set(theme_bw())

# Load soil data (found in /courses/EDS222/data on workbench)
soil_cores <- read_csv("data/exide-soil-cores.csv")
```

`sample()` draws samples from a vector (e.g., a column in a data frame)

```{r}
# Draws 50 random observations from the pb_ppm column
pb_ppm_sample <- sample(soil_cores$pb_ppm, size = 50)

# Visualize the whole distribution
ggplot(soil_cores, aes(pb_ppm)) +
    geom_histogram(binwidth = 20) +
    xlim(0, 1000)

# Visualize just our sample
tibble(pb_ppm = pb_ppm_sample) |> 
    ggplot(aes(pb_ppm)) +
    geom_histogram(binwidth = 50)
```

`slice_sample()` draws samples from a data frame instead.

```{r}
# Takes random samples from data frame, returns df
soil_sample <- soil_cores |> 
    slice_sample(n = 50)

ggplot(soil_sample, aes(pb_ppm)) +
    geom_histogram(binwidth = 50)
```

## Parametric Distributions

Visualize the PDF with `dnorm()` and other `d___()` functions.

```{r}
tibble(
    # Created to have the same mean and sd as empirical data above
    x = seq(-860, 1370, length.out = 1000),
    pdf = dnorm(x, mean = 254, sd = 371)
) |> 
    ggplot(aes(x,pdf)) +
    geom_line(linewidth = 1.5)
```

Draw random samples with `rnorm()`.

```{r}
normal_sample <- tibble(x = rnorm(25, mean = 254, sd = 371))
ggplot(normal_sample, aes(x)) +
    geom_histogram(binwidth = 100)
```

## Discussion Questions:

1. In the parametric distributions coding demo, I'm unclear about the `x = seq(-860, 1370, length.out = 1000)` part of the tibble creation code. How did you select those min and max values from looking at the earlier empirical example, and why did you specifically use `length.out` instead of `length` to specify the number of values in the sequence?